---
title: "Hayes_Metzger"
date: 2026-02-02
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
```

```{r}
x <-11

# Check if the value of x is greater than 10
if(x > 10)
{ 
  # Paste takes the value stored in x and combines that with a character string
  print(paste(x, "is greater than 10")) 
} else
{ 
  print(paste(x, "is less than or equal to 10")) 
}
```

#### Q1.1

```         
[1] "11 is greater than 10"
```

```{r eval = FALSE}
x <- "five"

# Check if the value of x is greater than 10
if(x > 10)
{ 
  # If x is > 10, multiple x by 2
  print(x*2)
} else
{ 
  # If x is not > 10, divide x by 2  
  print(x/2) 
}
```

```{r eval = FALSE, echo = FALSE}
traceback()
```

```{r}
# define a variable
x <- 11

# check the value of x using nested if-else statements
if (x < 10) {
  # if x is less than 10
  print("x is less than 10")
} else {
  # if x is exactly equal to 10
  if (x == 10) {
    print("x is 10!!!")
  } else {
    # if x is greater than 10
    print("x is greater than 10")
  }
}

```

```{r}
vec <- c(9, 10, 11, 12)
```

```{r}
# For 1 through the length of the vector "vec" 
for (i in 1:length(vec)) {
  
  # check the value of using nested if-else statements
  if (vec[i] < 10) {
    # if the element is less than 10
    print("value is less than 10")
  } else {
    # if the element is exactly equal to 10
    if (vec[i] == 10) {
      # if the element equals 10
      print("value is 10!!!")
    } else {
      # if the element is greater than 10
      print("value is greater than 10")
    }
  }
  
}

```

#### Q.1.2

```{r}
#First we're going to make a vector = y
y <- c(-2, 42, 0, 10)

# For 1 through the length of the vector "vec" 
for (i in 1:length(y)) {
  
  # check the value of using nested if-else statements
  if (y[i] < 0) {
    # if the element is less than 10
    print("value is Negative :(")
  } else {
    # if the element is exactly equal to 10
    if (y[i] == 0) {
      # if the element equals 10
      print("value is 0 my hero!!!")
    } else {
      # if the element is greater than 10
      print("value is 10 or the answer to life and the unviverse and everything #slay")
    }
  }
  
}
```

```{r}
library(lterdatasampler)
library(tidyverse)
```

#### Q1.3

Stress was measured by observing the amount of glucocorticoid metabolite present in pika feces. The stress variable is called "concentration_pg_m" and it's in picograms GCM/gram of dry pika feces.

#### Q1.4

A row represents a observation of individual pika poop, collected between June and September of 2018.

```{r}
# Make a histogram
nwt_pikas %>% 
  ggplot(aes(x = concentration_pg_g)) +
  # Add the histogram geom, which only needs an x-axis
  # Choose a binwidth of 500 picogram GCM/gram
  geom_histogram(binwidth = 500,
                 fill = "white",
                 color = "black")
```

```{r}
# Make a scatterplot with jittered points
nwt_pikas %>% 
  # We're adding a little placeholder axis just so we can see the point distribution
  ggplot(aes(x ="Placeholder", 
             y = concentration_pg_g)) +
  # Add the geom_jitter geom
  geom_jitter(width = 0.1)
```

```{r}
nwt_pikas_categ <- nwt_pikas %>% 
  # Call the new column stress_category
  mutate(stress_category = case_when(
    # When the value is > 5000, make the new column's value "Stressed!!!!"
    concentration_pg_g > 5000 ~ "Stressed!!!!!!",
    # Otherwise, make the new column's value "Chill"
    .default = "Chill"
  ))

# Check out the first 6 rows, but remove the utm columns just for visibility
head(nwt_pikas_categ %>% select(-c(utm_easting, utm_northing)))

```

#### Q1.5

```{r}
# Make a scatterplot with jittered points
nwt_pikas_categ %>% 
  # We're adding a little placeholder axis just so we can see the point distribution
  ggplot(aes(x ="Placeholder", 
             y = concentration_pg_g, 
             color = stress_category)) +
  # Add the geom_jitter geom
  geom_jitter(width = 0.1)
```

```{r}
nwt_pikas_categ2 <- nwt_pikas_categ %>% 
  # Create a new column called month
  # then, extract the month from the date using the month() function
  mutate(month = month(date)) %>% 
  # Lastly, relocate the month column after the date column so it's more easily visible to us
  relocate(month, .after = date)

head(nwt_pikas_categ2)
```

```{r}
nwt_pikas_summerstress <- nwt_pikas_categ2 %>% 
  mutate(summer_stress_category = case_when(
    (month == 6 | month == 7) & concentration_pg_g > 5000 ~ "Early summer stress",
    (month == 6 | month == 7) & concentration_pg_g <= 5000 ~ "Early summer chill",
    (month == 8 | month == 9) & concentration_pg_g > 5000 ~ "Late summer stress",
    (month == 8 | month == 9) & concentration_pg_g <= 5000 ~ "Late summer chill",
    .default = "NA"
  ))

head(nwt_pikas_summerstress)

```

```{r}
# Make a scatterplot with jittered points
nwt_pikas_summerstress %>% 
  # We're adding a little placeholder axis just so we can see the point distribution
  ggplot(aes(x = month, 
             y = concentration_pg_g,
             color = summer_stress_category)) +
  # Add the geom_jitter geom
  geom_jitter(width = 0.1)
```

#### Q2.1

We're using the palmerpenguins data set.

```{r}
#We are loading the library and then reading in penguins as penguins
library(palmerpenguins)
penguins <- penguins
```

#### Q2.2

For our for loop we want it to read the species column from penguins and return the unique species names in a vector. Then, the for loop is going to iterate the across the species names in the vector giving us a unique plot for each species.

#### Q2.3

```{r}
#| warning: false
#| message: false

# Store a vector of unique species names from the Species column of species in the penguins dataframe.
peng_names <- unique(penguins$species)

# We are making a for loop across the species names in the peng_names vetor
for (i in 1:length(peng_names)) {
  #this ine filters data for the interated species
  filt_data <- penguins %>% 
    filter(species == peng_names[i])
  
  # These ggpolt lines plot a scatter plot with a lm line.
  plot <- filt_data %>% 
    ggplot(aes(x = flipper_length_mm,
               y = body_mass_g)) +
    geom_point() +
    geom_smooth(method = "lm") +
    # This adds a title pasting in "Species and the interated species name
    ggtitle(paste("Species:", peng_names[i]))
  
  # this displays the plot(s)
  print(plot)
}
```

#### Q 2.4

If flipper length is \>=200, label as Big Ass Penguin and display points as different color in the ggplots

#### Q 2.5

```{r}
penguins <- penguins %>% 
  # Call the new column BAP
  mutate(BAP = case_when(
    # When the value is > 200, make the new column's value "BigAssPenguin"
    flipper_length_mm > 200 ~ "BigAssPenguin",
    # Otherwise, make the new column's value "Tinys"
    .default = "Tinys"
  ))
```

```{r}
# We are making a for loop across the species names in the peng_names vetor
for (i in 1:length(peng_names)) {
  #this ine filters data for the interated species
  filt_data <- penguins %>% 
    filter(species == peng_names[i])
  
  # These ggpolt lines plot a scatter plot with a lm line.
  plot <- filt_data %>% 
    ggplot(aes(x = flipper_length_mm,
               y = body_mass_g)) +
    #coloring points by BAP category
    geom_point(aes(color=BAP)) +
    geom_smooth(method = "lm") +
    # This adds a title pasting in "Species and the interated species name
    ggtitle(paste("Species:", peng_names[i]))
  
  # this displays the plot(s)
  print(plot)
}
```
